---
title: "EDA"
output: html_document
date: "2023-04-26"
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(knitr)

opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache=FALSE,
  include=TRUE,
  fig.width = 10, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```

```{r}
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyverse)
```

## Variable
- `Urban`: Urban/Rural Status (1 if Urban counties, 0 if Rural counties)
- `PhysicalHealth`: Number of days that participant's physical health is not good in the past 30 days
- `MentalHealth`: Number of days that participant mental health is not good in the past 30 days
- `PhysicalActivity`: participant who reported doing exercise in the past 30 days (1 if did, 0 if didn't)
- `Asthma`: Whether participants experienced asthma (1 if current, 2 if former, 3 if never)
- `Arthritis`: Whether participants diagnosed with some form of arthritis (1 if diagnosed, 0 if not diagnosed)
- `Sex`: (1 if male, 0 if female)
- `Race`: (1 if White, 2 if Black, 3 if American Indian or Alaskan Native, 4 if Asian, 5 if Native Hawaiian or other Pacific Islander only, 6 if other race, 7 if Multiracial, 8 if Hispanic)
- `Age`: (1 if Age 18-24, 2 if Age 25-34, 3 if Age 35-44, 4 if Age 45-54, 5 if Age 55-64, 6 if Age 65+)
- `BMI`: Calculated BMI based on Height and Weight. 
- `Education`: Level of education (1 if didn't graduate high school, 2 if graduated high school, 3 if attended college or technical school, 4 if graduated from college or technical school)
- `Smoke`: Levels of smoking (1 if smokes everyday, 2 if smokes some days, 3 if former smoker, 4 is never smoked)
- `Alcohol`: Number of alcoholic beverages consumed per week.
- `HHADULT`: Number of members in household
- `HIV`: Participants tested for HIV (1 if Yes, 0 if No)
- `Fruit`: Total fruits consumed per day
- `Vegetable`: Total vegetables consumed per day
- `Cholesterol`: Cholesterol is high (1 if Yes, 0 if No)
- `Stroke`: Had a stroke (1 if Yes, 0 if No)
- `MICHD`: Participants have reported having myocardial infarction(MI) or coronary heart disease(CHD) (1 if having, 0 if not having)

**Dependent variable**: `MICHD`

**Independent variables**: all other 

# Data Prep
```{r}
michd = read.csv("MICHD.csv")
str(michd)
```
```{r}
col_factor = c("Urban", "PhysicalActivity","Asthma","Arthritis","Sex","Race","Age",
              "Education","Alcohol", "Smoke","HIV","Cholesterol", "Stroke", "MICHD")
michd[col_factor] = lapply(michd[col_factor], factor)
str(michd)
```

```{r}
# Density Plot
michd.num <- michd %>% select(c("PhysicalHealth", "MentalHealth", "BMI", "HHADULT", "Fruit", "Vegetable", "MICHD"))

ggplot(melt(michd.num), aes(x = value, fill = MICHD)) +
  geom_density(alpha = 0.5) +
  facet_wrap( ~ variable, scale = "free")
```


```{r}
# Conditional density plots
a <- michd %>%
  ggplot() +
  geom_histogram(aes(x=PhysicalHealth,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('Number of Days Physical Health Not Good in past 30 Days') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))

b <- michd %>%
  ggplot() +
  geom_histogram(aes(x=MentalHealth,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('Number of Days Mental Health Not Good in past 30 Days') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))

c <- michd %>%
  ggplot() +
  geom_histogram(aes(x=BMI,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('BMI') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))



d <- michd %>%
  ggplot() +
  geom_histogram(aes(x=HHADULT,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('HHADULT') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))


e <- michd %>%
  ggplot() +
  geom_histogram(aes(x=Fruit,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('Total fruits consumed per day') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))

f <- michd %>%
  ggplot() +
  geom_histogram(aes(x=Vegetable,
                     fill=(MICHD==1)),
                 position='fill') +  
  theme_bw() +
  xlab('Total Vegatables consumed per day') +
  ylab('Frequency') +
  theme(axis.title=element_text(size=14), 
        axis.text=element_text(size=14), 
        legend.text=element_text(size=14)) +
  scale_fill_manual(name='', 
                    labels=c('No MICHD', 'Had MICHD'), 
                    values=c('grey', 'red'))


library(gridExtra)
grid.arrange(a,b,c,d,e,f, ncol = 2, nrow = 3)
```

# Correlation Plot
```{r}
# Select numeric variables
michd.cor = michd  %>% select(c("PhysicalHealth", "MentalHealth", 
                              "BMI", "HHADULT", "Fruit", "Vegetable"))

michd.cor$Age.rank <- rank(michd$Age)
michd.cor$Education.rank <- rank(michd$Education)
michd.cor$Smoke.rank <- rank(michd$Smoke)

c = cor(michd.cor, method = "spearman")

library(corrplot)
corrplot(c, addCoef.col = "red", tl.cex = 0.8)
```

# Clustering
```{r}
set.seed(123)
michd_matrix_c <- model.matrix(~ . - 1, data = michd[, colnames(michd)])

michd_matrix_c[,'BMI'] <- scale(michd_matrix_c[,'BMI'])
michd_matrix_c[,'HHADULT'] <- scale(michd_matrix_c[,'HHADULT'])
michd_matrix_c[,'Fruit'] <- scale(michd_matrix_c[,'Fruit'])
michd_matrix_c[,'Vegetable'] <- scale(michd_matrix_c[,'Vegetable'])
```

```{r}
set.seed(123)
twcv = function(k) kmeans(michd_matrix_c, k, nstart = 25)$tot.withinss
k <- 1:15
twcv_values <- sapply(k, twcv)
plot(twcv_values, type = 'b', pch = 19)
```

```{r}
library(factoextra)
k6 <- kmeans(michd_matrix_c, centers = 6, nstart = 25)
fviz_cluster(k6, geom = "point", data = michd_matrix_c)

k8 <- kmeans(michd_matrix_c, centers = 8, nstart = 25)
fviz_cluster(k8, geom = "point", data = michd_matrix_c)

k2 <- kmeans(michd_matrix_c, centers = 2, nstart = 25)
fviz_cluster(k2, geom = "point", data = michd_matrix_c)
```

# PCA
```{r}
pca = prcomp(michd_matrix_c)
fviz_pca_biplot(pca,repel = TRUE, col.var = "red", col.ind = "white", label = c("var"))
```
